---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# scheduler

Recurring calendar events are very common in everyday life. When they occur in
regular or quasi-regular patterns they are called 'schedules'. Schedules can and
do take many forms. While some are straightforward, others are complex 
making them difficult to conceptualise, use and program with.

`scheduler` is an R package the goal of which is to make dealing with
these schedules a little easier. It does this by providing a set of simple
to understand functions and a flexible framework for using them together. 
Together these allow users to compose more complex schedules from simple and 
intutive buildings blocks.

The package is in the early development stage and experimental. Users are 
cautioned against relying on it for anything of importance. It can be 
downloaded from Github:

```{r, eval = FALSE}
devtools::install_github("jameslairdsmith/scheduler")
```

