---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# scheduler
<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

Recurring calendar events are very common in everyday life. When they occur in
regular or quasi-regular patterns they are called 'schedules'. Schedules can and
do take many forms. While some are straightforward, others are complex 
making them difficult to conceptualise, use and program with.

`scheduler` is an R package the goal of which is to make dealing with
these schedules a little easier. It does this by providing a set of simple
to understand functions and a flexible framework for combining them. 
Together these allow users to compose more complex schedules from simple and 
intutive buildings blocks.

## Installation

The package is in the early development stage and experimental. Users are 
cautioned against relying on it for anything of importance. It can be 
downloaded from Github:

```{r, eval = FALSE}
devtools::install_github("jameslairdsmith/scheduler")
```

Feedback, bug report and suggestions are most welcome and can be made in the 
[issues tab](https://github.com/jameslairdsmith/scheduler/issues) of the Github
repo.

## Getting started

If you have not used `scheduler` before, the best place to get started is the
introductory vignette.
