---
title: "Holiday examples using `gs`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gs)
library(magrittr)
```

In this brief vignette I will give some examples to demonstrate how a set of 
national holidays can be put together using the `gs` grammar.

Using the grammar this way means that we can avoid the hassle of ennumerating
all the possible dates for national holidays and can rather rely on the 
pre-scripted logic to produce the apppriate dates for us when we request them.

In the future I plan to release a package that will catalogue this holiday
logic for a wide set of countries. The purpose of this vignette is only to 
provide some illustrative examples.

If you haven't heard of the `gs` package before or you are unsure what the 
grammar is, I recommend looking at the introductory vignette included in the
package.

# General

There are certain holidays which are common in many countries. One advantage
of the `gs` grammar is that the schedule objects only need to be created once
and can be re-used as necessary when they are required.

These common holidays are as follows:

```{r}
on_new_years_day <- on_yday(1)
on_christmas_day <- only_occur(on_mday(25), in_month("Dec"))
on_boxing_day <- only_occur(on_mday(26), in_month("Dec"))
```

## Easter

The date of Easter is notoriously tricky to calculate. For this reason there
is a shortcut function provided for it in `gs` called `on_easter_sunday()`.
This shortcut relies on the algorithm provided in the `timeDate` package.

```{r}
schedule_days(on_easter_sunday(), from = 2010, to = 2020)
```

There are also equivalent shortcut functions for `on_good_friday()` and
`on_easter_monday()` etc.

# United States Federal Holidays

In this next example we will build a schedule of the federal holidays in the
United States using the `gs` grammar. Each chunk of code below encodes the
rule of the holiday.

The `gs` grammer is designed to be readable, so you should be able to see what
the syntax is doing just by reading it. For example in the first chunk below
we are creating a schedule of Martin Luther King days, which in the US occur 
on the third Monday of January every year. To do this we find the intersection
(using `only_occur()`) of all days which are the third (using `on_third()`) 
Monday (using `on_wday("Mon")`) of every month with the schedule of days 
occuring in Janurary (using `in_month("Jan")`). The other holidays can be 
read similarly.

```{r}
on_mlk_day <- 
  on_third(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Jan"))

on_presidents_day <- 
  on_third(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Feb"))

on_memorial_day <-
  on_last(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("May"))

on_us_independence_day <- only_occur(on_mday(4), in_month("July"))

on_labor_day <-
  on_first(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Sep"))

on_columbus_day <-
  on_second(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Oct"))

on_veterans_day <- only_occur(on_mday(11), in_month("Nov"))

on_us_thanksgiving <- 
  on_fourth(on_wday("Thu"), within_given = "month") %>% 
  only_occur(in_month("Nov"))
```

To make these holidays into a single schedule, we combine them using the 
`also_occur()` function, which finds the union of two schedules. With the 
help of the pipe (`%>%`) operator, this forms a neat chain of operations.

```{r}
on_us_federal_holidays <-
  on_new_years_day %>% 
  also_occur(on_mlk_day) %>% 
  also_occur(on_presidents_day) %>% 
  also_occur(on_memorial_day) %>% 
  also_occur(on_us_independence_day) %>% 
  also_occur(on_labor_day) %>% 
  also_occur(on_columbus_day) %>% 
  also_occur(on_veterans_day) %>% 
  also_occur(on_us_thanksgiving) %>% 
  also_occur(on_christmas_day)
```

This schedule is then available for use. For example we can use the 
`schedule_days()` function to retrieve for us the occurances of US federal
holidays in 2019 and 2020.

```{r}
schedule_days(on_us_federal_holidays, from = 2019, to = 2020)
```

