---
title: "Holiday examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(gs)
library(magrittr)
```

This is a work in progress compilation of the various public holidays. 

Naturally I cannot ennumerate every public holiday in every country, but I 
hope the following examples demonstrate the process well enough. Hopefully
then users can fill in their own.

# General

```{r}
on_new_years_day <- on_yday(1)
on_christmas_day <- only_occur(on_mday(25), in_month("Dec"))
on_boxing_day <- only_occur(on_mday(26), in_month("Dec"))
```

## Easter

WIP

# UK

UK holidays vary by jurisdication. All the holidays are specified in the 
below chunk and are then assembled in the following sections.

```{r}
on_second_jan <- on_yday(2)

on_st_patricks_day <- only_occur(on_mday(17), in_month("Mar"))

on_early_may_bank_holiday <- 
  on_first(on_mday("Mon"), within_given = "month") %>% 
  only_occur(in_month("May"))

## ^ This needs some adjustment

on_liberation_day <- only_occur(on_mday(9), in_month("May"))

on_spring_bank_holiday <-
  on_last(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("May"))

on_june_bank_holiday <- 
  on_first(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("June"))

on_senior_race_day <-
  on_first(on_wday("Fri"), within_given = "month") %>% 
  only_occur(in_month("June"))

on_tynwald_day <- only_occur(on_mday(5), in_month("July"))
  
on_battle_of_the_boyne <- only_occur(on_mday(12), in_month("July"))

```

## England and Wales

```{r}
on_enland_wales_holidays <-
  on_new_years_day %>% 
  also_occur(on_early_may_bank_holiday)
```

## Scotland

## Northern Ireland

## Isle of Man

## Jersey and Guernsey

# United States Federal Holidays

```{r}
on_mlk_day <- 
  on_third(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Jan"))

on_presidents_day <- 
  on_third(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Feb"))

on_memorial_day <-
  on_last(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("May"))

on_us_independence_day <- only_occur(on_mday(4), in_month("July"))

on_labor_day <-
  on_first(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Sep"))

on_columbus_day <-
  on_second(on_wday("Mon"), within_given = "month") %>% 
  only_occur(in_month("Oct"))

on_veterans_day <- only_occur(on_mday(11), in_month("Nov"))

on_us_thanksgiving <- 
  on_fourth(on_wday("Thu"), within_given = "month") %>% 
  only_occur(in_month("Nov"))
```

```{r}
on_us_federal_holidays <-
  on_new_years_day %>% 
  also_occur(on_mlk_day) %>% 
  also_occur(on_presidents_day) %>% 
  also_occur(on_memorial_day) %>% 
  also_occur(on_us_independence_day) %>% 
  also_occur(on_labor_day) %>% 
  also_occur(on_columbus_day) %>% 
  also_occur(on_veterans_day) %>% 
  also_occur(on_us_thanksgiving) %>% 
  also_occur(on_christmas_day)
```



