---
title: "Working with recurring calendar events in R"
author: "James Laird-Smith"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Recurring calendar events are very common in everyday life. Some are simple
and therefore easy to program with. Others however are not and are difficult
to understand and exceeding difficult to program with.

The objective of `scheduler` is to make working with these recurring calendar
events a little easier. By providing an intutive set of building blocks and a
means of combining them together, users will be able to create and use 
schedules of arbitrary complexity.

To start with, we will need to laod the following packages:

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(lubridate)
library(scheduler)
library(magrittr)
```

`scheduler` builds atop the `lubridate` package, which provides various 
accessor functions for different elements of a date. For example, if I wanted
to know which day in a year a particlar date fell upon, lubridate provides a
function to access it, called `yday()`:

```{r}
lubridate::yday(as.Date("2000-01-01"))
lubridate::yday(as.Date("2000-12-31"))
```

`lubridate` provides a whole set of these accessor functions. For example, 
`mday()` to access the day of the month a date falls on, `wday()` for the day
of the week, `month()` for the month of the date, `year()` for the year of 
the date etc.

```{r}
first_day_millenium <- as.Date("2000-01-01")

lubridate::mday(first_day_millenium)
lubridate::wday(first_day_millenium)
lubridate::month(first_day_millenium)
```

`scheduler` fufils a different purpose, but uses the same naming scheme.
Whereas `lubridate` provides accessors to elements of a particualr date, the
purpose of `scheduler` is to let you specify the element of the schedule.
