---
title: "Work with recurring calendar events in R"
author: "James Laird-Smith"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning =  FALSE,
  message = FALSE
)
```

Recurring calendar events are very common in everyday life.  When a set of 
these events recur in a pattern, we call it a 'schedule'. While some 
schedules are simple, others can be difficult to work with. This is 
especially true when the pattern of events is irregular. The objective of 
`gs` is to provide a framework to make working with schedules simpler.

To start with, we will import `gs` along with the `magrittr` package, which
will be useful later on.

```{r setup}
library(gs)
library(magrittr)
```

`gs` builds atop the `lubridate` package and is designed for use
alongside it. For this reason, it is imported automatically when `gs` is
imported. When a `lubridate` function is used in this vignette, I will 
prefix it with `lubridate::` to make it clear where the function comes from. 
Users do not need to include the package name when writing their own code.

# Creating a schedule

In this section we will create a schedule of events occurring every year 
on New Year's Day (January 1st). This simple example which teach you the 
basics of schedules so that you can build-up to more involved examples later
on.

Ask yourself, what is unique about New Year's Day (from a calendar 
perspective)? It occurs on the first day of the month, but so do many other 
days. It occurs in January, but so do many other days. You would be right to
say that New Year's Day always occurs on the first day of the year. It is 
the only day that does so.

`gs` provides a function to create schedules where the events occur
only on certain days of the year. The function is called `on_yday()` and it 
accepts numeric input depending on which day of the year you wish to schedule
events for. For example, `on_yday(2)` would create a schedule of events 
occurring on the second day of every year. But in our case we want to create 
a schedule of events occurring on only the first day of every year. So we do
so as follows.

```{r}
on_yday(1)
```

You may be familiar with the `yday()` function from the `lubridate` package.
`lubridate` provides a set of accessor functions for getting the various
properties from a date or date-time object. The `yday()` function
returns whatever day of the year a date or datetime occurs on. For example:

```{r}
lubridate::yday(as.Date("2000-01-01"))
```

The equivalent function in `gs` is `on_yday()` and it does the opposite. 
Instead of the you providing a date and getting back the property of that date,
rather you give it the property you want and it will return a schedule of 
events with that particular property. So by computing `on_yday(1)` above, I 
created a schedule of all the days occurring on the first day of the year.

By itself, this isn't that useful. What will help make it more so is that 
we can make this schedule into an object like so:

```{r}
on_new_years_day <- on_yday(1)
```

We now have the `on_new_years_day` schedule object that we can put to use. You 
can think of this schedule as encompassing all the possible occurrences of
New Year's Day throughout time. It is not limited to just one particular date
or dates.

# Using schedules

Now that we have a schedule object, we can begin using it.

## Testing for events

The first thing we can use schedules for is to find out whether certain dates
fall on them or not. Lets create an arbitrary set of dates using base R and 
call them `my_dates`:

```{r}
my_dates <- seq.Date(from = as.Date("2001-01-01"),
                     to = as.Date("2001-01-10"),
                     by = "1 day")

my_dates
```

We can then use the `happen()` function provided by `gs` to find out which of 
these dates falls on New Year's Day. `happen()` takes a schedule object as 
its first argument and a date or vector of dates as the second argument. The 
idea is that the syntax is readable. We are asking for the events which 
'happen',  'on New Year's day' from '`my_dates`':

```{r}
happen(on_new_years_day, my_dates)
```

We can see from this that only the first of our dates fall on New Year's Day.
In this example you can validate this simply by inspecting the dates.

Note that you do not have to make the `on_new_years_day` schedule into an 
object if you don't want to. We could have achieved the same result as follows:

```{r}
happen(on_yday(1), my_dates)
```

But, for reasons that will become clear, it is much more useful to create 
schedules objects. Whenever you do so, I also recommend giving them a 
descriptive name and starting them with the prefix 'on_' or 'in_'. Among other
things, this will make the syntax natural and readable.

## Getting events

The next useful thing we can do with a schedule object is get the events from 
it. This is done using the `schedule_days()` function, which accepts a 
schedule as its first argument.

One snag with doing this however is that the total number of New Year's Days is 
theoretically infinite. So if we try and run `schedule_days(on_new_years_day)`
we will get an error. What we can do is place date limits on the 
resulting output using the `from` and `to` arguments:

```{r}
schedule_days(on_new_years_day, 
              from = as.Date("2001-01-01"), 
              to = as.Date("2001-01-10"))
```

This gives us all the occurrences of New Year's Day within the boundaries we
have specified. Let's expand these boundaries so we get more occurrences of
New Year's Day. Here I extend the boundaries from the start of 1995 to the 
end of 2005.

```{r}
schedule_days(on_new_years_day, 
              from = as.Date("1995-01-01"), 
              to = as.Date("2005-12-31"))
```

Because this is a little cumbersome to type, `schedule_days()` allows you to
abbreviate `to` and `from` to only the numeric years you wish to use as 
boundaries. These go from the start of the `from` year to the end of the `to`
year. That means this code is equivalent to what we just ran:

```{r}
schedule_days(on_new_years_day, from = 1995, to = 2005)
```

If the events we want only occur in a single year, we can use the `during`
argument as a further numeric shortcut, which is the equivalent of setting `to`
and `from` to the start and end of one particular year:

```{r}
schedule_days(on_new_years_day, during = 2005)
```

# Further basic schedules

The `on_yday()` function isn't the only type of schedule that you can create.
`gs` provides a whole host of functions to create schedules. These all follow
from the syntax of the accessor functions found in `lubridate`. In each case
the same function text name is prefixed by either `on_` or `in_` depending on
how one would naturally say it.

For example, the `on_mday()` function creates a schedule of events occurring on
the specified days of every month. Imagine your job pays your salary 
on the 25th of each month; if you wanted to create a schedule of your paydays
you would do so as follows:

```{r}
on_payday <- on_mday(25)
```

If instead you get paid weekly every Friday, you can use the `on_wday()` 
function which accepts either a day number (where 1 is Sunday by default), a
day name or a day abbreviation as the first argument. This means each of these
are equivalent and create the schedule of your weekly paydays:

```{r}
on_weekly_payday <- on_wday(6)
on_weekly_payday <- on_wday("Friday")
on_weekly_payday <- on_wday("Fri")
```

I won't go over all the available functions here. You can see them for 
yourself in the package reference documentation. They all follow the same
principle of generating schedules from the inputs you provide to them. Some
will also appear in the upcoming examples.

Each of the functions will accept multiple inputs which in each case creates
a schedule of events occurring on both the inputs. For example 
`on_wday("Mon", "Wed", "Fri")` creates a schedule of events occurring on all
of those days of the week. For convenience the `on_weekday()` and `on_weekend()` 
functions are also provided in the package without you having to create them.

# Joining schedules

By itself, New Year's Day isn't that interesting (from a calendar perspective).
We could have achieved the same thing using base R. The power of `gs` comes
when dealing with more intricate schedules. For this, we need something more.
`gs` allows you to compose more complex schedules by combining basic ones.

## Intersecting schedules

Say that, instead of New Year's Day, we were interested in Christmas day. Ask
yourself, what is uniquely special about Christmas day (from a calendar 
perspective)? You may think of using `on_yday()` again and say that Christmas 
occurs on the 359th day of the year and so create the schedule `on_yday(359)`.
But this wouldn't work because in a leap year Christmas day is the 360th day of
the year. The only way to create a schedule of events on Christmas is to
schedule them on December 25th.

`gs` provides some functions to help us along the way. We have the `in_month()`
function which creates a schedule of events in a given month. We could use that,
but `in_month(12)` would give us all the events in December and we are only 
interested in the 25th one. We could also use the `on_mday()` function. But
`on_mday(25)` is going to give us a schedule of events occurring on the 25th of
every month, not just December.

What we need is the **intersection** of these two schedules. For this purpose
`gs` provides the `only_occur()` function, which accepts two schedules and 
returns a single schedule which is the intersection of the two inputs.

```{r}
on_christmas_day <- only_occur(on_mday(25), in_month(12))
```

Again the syntax is designed to be readable. We are creating a schedule called
`on_christmas_day` which 'only occur' on 25th day of the month and 'only occur'
in the twelfth month of the year.

Now that we have this new `on_christmas_day` object, we can do with it the 
same set of things we did with the `on_new_years_day` object. We can test 
if certain dates fall upon it with the `happen()` function. Or we can get 
the events from it using the `schedule_days()` function:

```{r}
schedule_days(on_christmas_day, from = 2000, to = 2004)
```

If I wanted to create a schedule of Boxing Day days, the process is the same.
Boxing day occurs on Dec 26th.

```{r}
on_boxing_day <- only_occur(on_mday(26), in_month(12))

schedule_days(on_boxing_day, from = 2000, to = 2004)
```

## Uniting schedules

Sometimes, instead of finding the intersection between schedules, you will 
instead want to unite schedules. For example, say now that we have the schedules
for `on_new_years_day` and `on_christmas` we wanted to create a 
schedule of public holidays. To do this, `gs` provides the `also_occur()`
function which takes two schedules as input and returns a single unified 
schedule as its output.

```{r}
on_public_holidays <- also_occur(on_christmas_day, on_new_years_day)

schedule_days(on_public_holidays, from = 2000, to = 2004)
```

We now have a new schedule of public holidays available for us to use further.

One other thing to take note of is that the schedules we made use of in this
step were themselves generated in earlier steps. This can be done an 
arbitrary number of times and allows huge flexibility when creating your own 
schedules.

The syntax of the `also_occur()` and `only_occur()` functions also lend 
themselves to piping (`%>%`) using the `magrittr` package. For example, if I 
wanted to include Boxing Days in my public holiday schedule I could 
re-make it as follows:

```{r}
on_public_holidays <-
  on_christmas_day %>% 
  also_occur(on_new_years_day) %>% 
  also_occur(on_boxing_day)

schedule_days(on_public_holidays, from = 2000, to = 2004)
```

Again the syntax is readable. We are creating a public holidays schedule, where
the events occur on Christmas day, and 'also occur' on New Year's Day and 
'also occur' on Boxing Day.

## Inverting schedules

Once you have a schedule it is often useful to have way to invert it so as to
get the events that do not occur on that schedule. This can be done using
the `doesnt_occur()` function. For example, say that 
instead of focusing on holidays, you instead wanted to get a schedule of 
working (or business) days. These are the days that occur on neither weekends 
nor on public holidays. You could get that schedule of events as follows. First
you would create the schedule of non-working days as follows:

```{r}
on_non_working_days <- 
  on_weekend() %>% 
  also_occur(on_public_holidays)
```

And then invert it using `doesnt_occur()`:

```{r}
on_business_days <- doesnt_occur(on_non_working_days)
```

You then have a new schedule of business days to work with that can be used in
exactly the same ways already shown.

# Nth occurrences within periods

One of the most common schedule patterns are when events take place on the nth
occurrence of another schedule within a given period. For example, in the 
United States, Martin Luther King Jr. Day is celebrated on the third Monday in
January.

This date rule has a few components. 'Monday', 'January' and 'third occurrence
within a given month'. Mondays don't have their own schedule, 
but instead are created by `on_wday("Mon")`. January also doesn't have it's 
own schedule, but is instead created using `in_month("Jan")`.

```{r}
in_jan <- in_month("Jan")
on_monday <- on_wday("Mon")
```

To find the third occurrence of a scheduled event within a period `gs` provides 
the `on_third()` function, which accepts a schedule as its first argument and 
a period type string as its second argument (called `within_given =`). So the
schedule of every third Monday of the month is created as follows:

```{r}
on_third_monday_month <- on_third(on_monday, within_given = "month")
```

But because Martin Luther King Jr. Day doesn't occur every month, it only 
occurs in January, we have to refine our schedule one step further:

```{r}
on_mlk_jr_day <- 
  on_third_monday_month %>% 
  only_occur(in_jan)
```

We now have our schedule can can call on it for our purposes as before:

```{r}
schedule_days(on_mlk_jr_day, from = 2010, to = 2020)
```

`on_third()` isn't the only function provided for this purpose. There is also
`on_first()`, `on_second()`, `on_fourth()` and `on_last()`. These are all 
convenience functions of `on_nth()` which instead accepts an integer as its
first argument. Positive integers encode the nth occurring event within the 
period; negative integers encode the nth last occurring events of the period.

As a somewhat bizarre example, say that you play for a football team whose
matches are on the 12th either Monday, Wednesday or Friday of every quarter. 
You could find your match days as follows:

```{r}
on_monday_wednesday_friday <- on_wday("Mon", "Wed", "Fri")

on_football_days <- on_nth(12, 
                           on_monday_wednesday_friday, 
                           within_given = "quarter")

schedule_days(on_football_days, from = 2019, to = 2020)
```


# Conclusion

This vignette has provided an introduction to the `gs` package, which 
implements a grammar of recurring calendar events in R. 

`gs` has greater capabilities than just those explained above. The remainder 
of the package documentation provides greater detail on both the features 
shown and those I have not had time to mention.

Because `gs` is a grammar, it provides the building blocks for you as a user
to combine the different parts in new and interesting ways and to make
your own schedules based on your own needs.
