---
title: "understanding-schedule-limits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{understanding-schedule-limits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Once you have created a schedule object with your desired specifications, 
`scheduler` allows you to retrieve the dates which fall on that schedule. This
is done using the `schedule_*()` family of functions. 

Under normal circumstances this process should be seamless and intuitive, but
for those looking for a deeper understanding of how it works. To do so, you
need first to load the package:

```{r setup}
library(scheduler)
```


The nature of recurring calendar events is that there are a theoretically 
infinite number of possible occurrences. If you ask `scheduler` to retrieve
for you all the occurrences of Christmas, you will see that it refuses to do so:

```{r}
on_mday(25) %>% 
  only_occuring(in_month("Dec")) %>% 
  schedule()
```

This is possible because `scheduler` knows the number of possible Christmases
is infinite and that it cannot produce for you the complete set. To make your
request possible, you must place some limit on which Christmases you wish to
schedule.
